---
title: "Introduction to API's"
output: html_notebook
---

# Working with API with no key required

-The main goal here is to learn how to convert JSOn into a table.
-To do that we require to essential packages 
```{r}
library(httr)
library(jsonlite)

```

## Create a GET response to call the API
```{r}
foodbank<-GET('https://www.givefood.org.uk/api/2/foodbanks/')

# View  API GET response results
foodbank$status_code  # 200 
str(foodbank)

# View the main source of the data needed from the API - Normally content
str(foodbank$content)
```
🔑 **Note** Always check the status code which should be 200 to show the data has been extracted succesfully


## Convert the raw content to text (In this case, a JSON string)
```{r}
foodbankcontent <-httr::content(foodbank,as ='text')

# View JSON string results
str(foodbankcontent)
```
## Convert JSON string to a dataframe and view data in a table format
```{r}
foodbankJSON <- jsonlite::fromJSON(foodbankcontent)

View(foodbankJSON)
```


# Working with API that requires one to have a key
- Remember the essentials packages we will rely on is still jsonlite and httr
>
The data we are going to extract using the API key Reed.co.uk which is a site in UK that gives you all the information about all the jobs they have in their website

```{r}
# Create a GET response to call the API's
# For this job search we would like to filter our data for the following information
#  1. Analyst - From the job description
#  2. London - The location of the job 
#  3. Distance 15 km - The job place is 15 km within London

jobsearch<- GET('https://www.reed.co.uk/api/1.0/search?keywords=analyst&location=london&distancefromlocation=15?details=true',  #?details=true 💡
                httr::authenticate( # For API's with authentication
                  user = "e9914e3a-0d2d-41ec-8b89-19d1473b5419",  # REED_API_Key
                  password = "",
                  type = "basic"))
# View API GET response results
jobsearch$status_code
str(jobsearch)

# View the main source of the data needed from the API -Normally content
str(jobsearch$content)

```


💡 **Note** The last part of the https we included `?details=true` this bit of logic that would set what that particular url data will be  

## Hidden keys
-In the case where you want to have your API hidden you can the following package to read it in to get request
```{r}
library(dotenv)
load_dot_env("cred_env")  # cred_env is the name given to the file containing the API key am guessing it should be within your working directory
```


## Convert the raw content to text 
```{r}
jobsearchcontent<- content(jobsearch,as = 'text')

# View JSON string results
str(jobsearchcontent)
```

## Convert the JSON string to a dataframe and view the data in a table
```{r}
jobsearchJSON<-jsonlite::fromJSON(jobsearchcontent)

View(jobsearchJSON$results)

```


# Fuel economy API tutorial

## Example 1 : Fuel prices
```{r}
# Build urls
base_url<-'www.fueleconomy.gov/ws/rest/' #Every API request is gonna involve varies for different API
info_url<-'fuelprices'
full_url<-paste0(base_url,info_url)
full_url
```


## Call API
```{r}
api_call<-GET(full_url)
```


## API response
```{r}
api_call$status_code
# 200 indicates the api fetched the data requested

api_call$content

```


## Convert response to usable data
```{r}
api_char<-rawToChar(api_call$content)

api_JSON<-fromJSON(api_char,flatten = TRUE)
api_JSON

api_JSON$diesel
```

## EXAMPLE 2: Vehicle Information 
```{r}
# Build urls
base_url

new_info_url<-"vehicle/menu/options?year=2012&make=Honda&model=Fit"

new_full_url<-paste0(base_url,new_info_url)
new_full_url
```

## Call API
```{r}
api_call<-GET(new_full_url)
```

## Check response
```{r}
api_call$status_code

```
# Convert the JSON file to a data table
```{r}
# Option 1
AB<-content(api_call,as ='text')

ABC<-fromJSON(AB)
ABC$menuItem

# Option 2
api_char<-rawToChar(api_call$content)

api_JSON<-fromJSON(api_char,flatten = TRUE)

df<-api_JSON$menuItem
df
```
































































