---
title: "Stock Analysis"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

# References 

* [Chi-square calculator]('https://www.socscistatistics.com/pvalues/chidistribution.aspx')

```{r}
# GLM's
# R_squared
R_squared<-function(x){
  'x must be a glm model'
  null.dev=x$null.deviance
  resid.dev=x$deviance
  return(1-(resid.dev/null.dev))
}

# Chi square value for the GLM's while looking for the goodness of fit in the model
Chisq_GLM=function(x){
  'x must be a glm model'
  null.dev=x$null.deviance
  resid.dev=x$deviance
  return(null.dev-resid.dev)
}
  
```


```{r}
library(tidyverse)
```

This project was started to keep track of the day to day analysis of the
hotel and try to come up with a continuous update of the data

## Import 2023 data

> Use the previous merged data

# Import 2024 data

> Try to use the key raw materials to predict the sales : 1)Ngano
> 2)Sugar 3)Ugali 4)Milk 5)Meat 6)Matumbo 7)Managu 8)Rice 9)Kuku 10)Food
> remained(Kuku,Beef,Ugali) 11)Sales

```{r}
library(readxl)
# Jan 
Jan <- read_excel("Hotel_24.xlsx", sheet = "January ")
# Feb 
Feb <- read_excel("Hotel_24.xlsx", sheet = "February")

# March 
March <- read_excel("Hotel_24.xlsx", sheet = "March")

# April
April <- read_excel("Hotel_24.xlsx", sheet = "April")

#May

May <- read_excel("Hotel_24.xlsx", sheet = "May")
```

# Test data

```{r}
# Jan-March
Jan %>% select(Ngano:Rice,Amount)->A
Feb %>% select(Ngano:Rice,Amount)->B
March %>% select(Ngano:Rice,Amount)->C

```

# Train data

```{r}

#April-May
April %>% select(Ngano_Chapo:Rice,Amount)
May %>% select(Ngano_Chapo:Rice,Amount)

```

# Combine the Ngano variables

```{r}
#April
April %>% select(starts_with('Ngano')) %>% rowSums(na.rm = T)->Ngano
April %>% select(Ugali:Remaining_Food,Amount) %>% cbind(Ngano)->April_N

# May
May %>% select(starts_with('Ngano')) %>% rowSums(na.rm = T)->Ngano
May %>% select(Ugali:Remaining_Food,Amount) %>% cbind(Ngano)->May_N

```

# Merge April-May datset

```{r}
rbind(April_N,May_N)->Merged_data

Merged_data %>% select(-Remaining_Food)->data_x

colSums(is.na(data_x))

```

# Merge Jan- March data to data_X(April-May)

```{r}
rbind(A,B,C,data_x) %>% drop_na()->Stock_data
Stock_data

```

# Create a GLM model for the stock data(Use as the base model)

```{r}
Stock_data

# Check properties of the dependent variable
Stock_data$Amount

shapiro.test(Stock_data$Amount)  # H0: Their is  normality in the data

ks.test(Stock_data$Amount,'pnorm')

# The Amount variable is not normal distributed thus a glm model is best preffered 
plot(density(Stock_data$Amount))
ggplot2::ggplot(data = Stock_data,aes(Amount))+
  geom_density()+
  geom_vline(aes(xintercept = mean(Amount),linetype = "Mean"),
             col ="blue")+
  geom_vline(aes(xintercept = median(Amount), linetype="Median"),
             col = "red")+
  theme_classic()
# The data is rightly  skewed
```

# Gamma regression

```{r}
# Canonical Form
fit.gam=glm(Amount~.,data = Stock_data,family = Gamma)
summary.glm(fit.gam)

# Log-Link
fit.gam.log=glm(Amount~.,data = Stock_data,family = Gamma(link = 'log'))
summary.glm(fit.gam.log)
```

# Inverse-gaussian

```{r}
# Inverse
fit.ig.inv=glm(Amount~.,data = Stock_data,family = inverse.gaussian(link = 'inverse'))
summary.glm(fit.ig.inv)
# Canonical
fit.ig.invquad=glm(Amount~.,data = Stock_data,family = inverse.gaussian)
summary.glm(fit.ig.invquad)

# Number of predictors is 9
# To compute the chi square to test the model fitness
0.00030818-0.00021302 #[9.516e-05] with a degree of freedom of 9
chisq.test(fit.ig.inv)
43.230-16.713
```

# Findings

* Performing a chi-square test to check for the goodness of fit in the data
* Using ANOVA to identify the model selection

```{r}
# Chisq test


model_fit=glm(Amount~Milk+Kuku+Meat,data = Stock_data,family=inverse.gaussian(link = 'inverse'))

summary.glm(model_fit)

0.00030818-0.00022396
#The P-Value is 1. The result is not significant at p < .05.
# The model performs bad


# Fitting using Gamma glm using it canonical form (inverse)
model_fit2=glm(Amount~Kuku+Meat,data = Stock_data, 
               family = Gamma()
               )
summary.glm(model_fit2)

model_fit2$null.deviance-model_fit2$deviance
# Computing the goodness of fit
4.1855-3.1197    #[1.0658] with a degree freedom of 2
#The P-Value is .5869. The result is not significant at p < .05.
R_squared=1-model_fit2$deviance/model_fit2$null.deviance ;R_squared
#[1] 0.2546332
#In this case the deviance is(residual deviance)

# Interaction effect
model_fit3=glm(Amount~Kuku*Meat,family = Gamma,data = Stock_data)
summary.glm(model_fit3)

Chisq_GLM(model_fit3) # 1.090872 with df=3
#The P-Value is .779295. The result is not significant at p < .05.

model_fit4=glm(Amount~Kuku*Meat,family = inverse.gaussian(link = 'inverse'),data = Stock_data)
summary.glm(model_fit4)
Chisq_GLM(model_fit4) #8.166275e-05
#The P-Value is .9999 The result is not significant at p < .05.

anova(model_fit,test = 'Chisq')
anova(model_fit2,test = 'Chisq')
anova(model_fit3,test = 'Chisq')
anova(model_fit4,test = 'Chisq')
anova(fit.ig.inv,test = 'Chisq')

# Model Selection
summary(MASS::stepAIC(fit.ig.inv,trace = 0))
summary(MASS::stepAIC(fit.gam,trace = 0))

Best_model<-glm(formula = Amount ~ Milk + Kuku + Meat + Rice, 
    family = inverse.gaussian(link = "inverse"),data = Stock_data)
summary(Best_model)

anova(Best_model,test = 'Chisq')
anova(Best_model,model_fit4,model_fit3,test = 'Chisq')

# Computing various r_squared for the model 
R_squared(Best_model)  #0.2927708
R_squared(model_fit4) #0.2649873
R_squared(model_fit2) #0.2546332
R_squared(model_fit3) #0.2606302
R_squared(fit.ig.inv) #0.313553
R_squared(fit.gam) #0.3113659

# 
```
💡 **Note**

However, this R2 has a different interpretation than in linear regression. In particular:
- It is not the percentage of variance explained by the model, but rather a ratio indicating how close is the fit to being perfect or the worst.
- It is not related to any correlation coefficient.

## Visualization

```{r}
plot(fit.gam,1)
plot(fit.ig.inv,1)
plot(fit.ig.invquad,1)
```

#



```{r}
anova(model_fit2)
model_fit2
write.csv(Stock_data,"Hotel_24.csv")
```




# Assumption
```{r}

# 1.Multicollinearity
Stock_data





```

