---
title: "R Notebook"
output: html_notebook
---

## Import data 
```{r}
May_data<-read_csv('D:\\Hotel\\Hotel_24 - May.csv') #Up to date 28
May_data
```



## Prepare data
```{r}
# Convert char to factors and Dates to their format
May_data %>% mutate(Remaining_Food=as.factor(Remaining_Food),
                    Date=as.Date(Date,tryFormats = "%d/%m/%Y"))->Clean_data
# Check for missing data
colSums(is.na(Clean_data)) # No missing data

# Check for duplicates
sum(duplicated(Clean_data)) # No duplicates
```



```{r}
#
Clean_data %>% select(contains('Man'),-Managu) %>% rowSums()->Managu

Clean_data %>% select(-Managu) %>% mutate(Managu=Managu,Managu_F=Managu) %>% mutate(Managu=case_when(Managu <=4 ~ '1',
       Managu>4 & Managu<=8 ~'2',
       Managu>8 & Managu<=12 ~'3',
       Managu>12 & Managu<=16 ~'4',
       Managu>16  ~'5',),Managu=as.factor(Managu))->Clean_data2

#
Clean_data2 %>% select(Amount:Managu_F,Soda,Dasani_.5ltr:Water_1ltr,Eggs)->train_data;train_data
train_data %>% select(contains('Ngano')) %>% rowSums()->Ngano
train_data %>% select(-contains('Ngano')) %>% mutate(Ngano=Ngano)->Train_data
```



## Menu Engineering

For this section we are going to split the data to be able to perform menu engineering on the various dishes as well as snacks
To start of we are going to work  with the snacks section 

## 1. Snacks
```{r}
# snacks section
Clean_data %>% select(chapo:`Sausage/Smokie`)-> Snacks

Snacks %>% colSums()

# chapo          Ndazi             Tm           cake          Hcake 
#  1325            994            132              0            115 
#  Eggs         Omelet Sausage/Smokie 
#   246              2             20 
```



## 2.Mix
```{r}
Clean_data %>% select(ChapoMix:PilauMix)->Mix
Mix
colSums(Mix)
# ChapoMix  Walimix Ugalimix PilauMix 
#     129      134      147       45 

# Check normality
shapiro.test(Mix$Walimix)  # p-value = 0.2583
shapiro.test(Mix$Ugalimix)  #p-value = 0.1131

# Check equality of variances

# Perform t-test
t.test(Mix$Ugalimix,Mix$Walimix)
t.test(Mix$Ugalimix,Mix$ChapoMix)

## Plots
Count<-colSums(Mix)
Data_M<-as.data.frame(Count)
Data_M
 
129+134+147+45

```

## Beef
```{r}
Clean_data %>% select(BeefChapo:BeefPilau)->Beef
Beef
# BeefChapo BeefUgali  BeefRice BeefPilau 
#       32        96        23        15 

CountBeef<-colSums(Beef)
Data_B<-as.data.frame(CountBeef)
Data_B

32  +      96   +     23     +   15
```


## Kuku
```{r}
Clean_data %>% select(KukuChapo:KukuPilau)->KUKU
KUKU
# KukuChapo KukuUgali  KukuRice KukuPilau 
#        76       130        29         3 
CountKuku<-colSums(KUKU)
Data_K<-as.data.frame(CountKuku)
Data_K

 76   +    130   +     29     +    3 
```
## Managu
```{r}
Clean_data %>% select(UgaliManagu:PilauManagu)->Managu
Managu
# UgaliManagu  RiceManagu ChapoManagu PilauManagu 
#          87           1           10           0 
CountManagu<-colSums(Managu)
Data_man<-as.data.frame(CountManagu)
Data_man

 87     +      1      +     10     +      0 
```

## Drinks

```{r}
Clean_data %>% select(Tea:Water_1ltr)->Drinks

colSums(Drinks)

#        Tea  Black Coffee   WhiteCofee     Lemon Tea    Concusion 
#        3300           69           82            0            0 
#        Predator    Soda  Plastic Soda    Dasani_.5ltr  Dasani_1ltr 
#           7          202            0           16           13 
#     Water_.5ltr   Water_1ltr 
#          33           21 
```





## Model 

Am interested to understand the relationship between beef and kuku on the day to day sales

```{r}
 shapiro.test(Clean_data$Amount)
# Data appears to be normaly distributed
summary(Clean_data$Amount)

ggplot(data = Clean_data,aes(x=Amount))+geom_density()+
  geom_vline(aes(xintercept = mean(Amount),linetype='Mean'),col ='red')+
  geom_vline(aes(xintercept = median(Amount),linetype='Median'),col ='blue')


summary.lm(lm(Amount~Meat+Kuku+Milk+Ngano_Chapo-1,data = Clean_data))


# Summing up the ngano variables
summary.lm(lm(Amount~Meat+Kuku+Milk+Ngano-1+Soda,data = Train_data))
# Odd the no. ngano used for the raw materials have no significant as compared to chapo alone.

# Best model
summary.lm(lm(Amount~Meat+Kuku+Milk+Ngano_Chapo-1+Soda+Eggs,data = Clean_data))

summary.lm(lm(Amount~Meat+Kuku+Milk+Ngano_Chapo+Soda+Eggs+Remaining_Food,data = Clean_data))

summary.lm(lm(Amount~Meat+Kuku+Milk+Ngano_Chapo-1+Soda+Eggs+Water_.5ltr,data = Clean_data))

summary.lm(lm(Amount~Meat+Kuku+Milk+Ngano_Chapo-1+Soda+Eggs+Water_.5ltr+Ngano_ndazi,data = Clean_data))

summary.lm(lm(Amount~Meat+Kuku+Milk+Ngano_Chapo-1+Soda+Eggs+Water_.5ltr+Dasani_.5ltr,data = Clean_data))

Model_best<-lm(Amount~Meat+Kuku+Milk+Ngano_Chapo-1+Soda+Eggs,data = Clean_data)

# Check assumptions if are violated

library(performance)

check_collinearity(Model_best)  # High correlation Meat, Milk

check_heteroscedasticity(Model_best) # Not violated

check_autocorrelation(Model_best) # Not violated

shapiro.test(Model_best$residuals) # Not violated

# Check linearity
check_model(Model_best)  # Not violated


```

From the may model data the most significant variables in the model are:
 - Meat
 - Kuku
 - Milk
 - Ngano_chapo
 - Soda
 - Eggs
 
with coefficient of meat with the most significant impact on the model

From the assumption section investigate how to prevent multicollinearity in the data using a lasso-ridge elastic net regression model on the data

## Assignment use the previous months data to understand how the model is performing on the unsupervised learning

```{r}
# Test data we will use the March dataset

March_data<-read_csv('D:\\Hotel\\Hotel_24 - March.csv')

March_data %>% select(Ngano,Meat,Kuku,Milk,Soda,Eggs) %>% rename(Ngano_Chapo=Ngano)->Test_data

March_data %>% select(Amount)->Test_labels

Pred_labels<-predict(Model_best,Test_data)

cbind(Test_labels,Pred_labels)

# Mae
sum(abs(Pred_labels-Test_labels))/length(Pred_labels) #2798.374
```

**Note** The models performs poorly at predicting the test data this is probably caused by over fitting on the test data.

**Assignment**  Find remedies to remove over fitting in models


## *Overfitting in models*

Statisticians have conducted simulation studies which indicate you should have at least 10-15 observations for each term in a linear model. The number of terms in a model is the sum of all the independent variables, their interactions, and polynomial terms to model curvature.

For instance, if the regression model has two independent variables and their interaction term, you have three terms and need 30-45 observations. Although, if the model has multicollinearity or if the effect size is small, you might need more observations.



## How to detect overfitting in models

However, for linear regression, there is an excellent accelerated cross-validation method called predicted R-squared. This method doesn’t require you to collect a separate sample or partition your data, and you can obtain the cross-validated results as you fit the model. Statistical software calculates *predicted R-squared* using the following automated procedure:



### Predicted r_squared function
```{r}
Pred_Rsq<-function(pred_labels,test_labels){
  # Find the SST ,SSR ,SSE
  # Calculate the r-Squared for the predicted observations
  
  
  SST<-sum((test_labels-colMeans(test_labels))^2)
  SSE<-sum((test_labels-pred_labels)^2)
  
  R<-(1-(SSE/SST))
  return(paste0('Predicted R_squared : ',round(R,3),'%'))
  #return(R)
}

colMeans(Test_labels,na.rm = T)
mean(Pred_labels)
SST<-sum((Test_labels-colMeans(Test_labels))^2)
SSE<-sum((Test_labels-Pred_labels)^2)
SSE;SST
1-(SSE/SST)

# Test it on the Model_best predictions
Pred_Rsq(Pred_labels,test_labels = Test_labels)
```


## Food for thought 

Try using the snacks variable(Do not use the stock data ) in predicting the daily sales to understand which models best works with the data
```{r}
Clean_data2 %>% dplyr::select(chapo:Tm,Hcake,Eggs,Tea:WhiteCofee,Soda,Dasani_.5ltr:Water_1ltr,Amount,Ugali:Managu,-Managu,-Matumbo)->New_data


summary.glm(glm(Amount~.,family = Gamma(link = 'log'),data = New_data))
summary.glm(glm(Amount~Meat+Kuku+Tea+Eggs,family = Gamma(link = 'log'),data = New_data))

```

```{r}
Test_data

Train_data
```






